Topics Causal Inference

### 2024-08-15 documents/paper.tex
Fix latex compilation.
Add to paper.

@@

\title{Topics Causal Inference}
\thanks{Bonn University. Email: s93jbudd[at]uni-bonn.de}

\author{Julian Budde}

@@

\paragraph*{Approximations to the CATE}
Another approach is taken in~\cite{semenova2021debiased}:
They propose to focus on a \textit{low-dimensional} subset $X$ of the covariates $Z$, which the researcher has to decide on ex ante.
Then, the new target for inference becomes the \textit{partial} CATE $s_0(X)$.
Their paper then simultaneously allows for \textit{unconfoundedness} which poses the challenge of estimating $p(Z)$ --- potentially using machine learning methods --- in a first step.
In a second step, they the nfocus on a \textit{linear} approximation to the (partial) CATE\@.
Hence, they essentially use the following approximations to make the problem solvable:
\begin{equation*}
    s_0(Z) \approx s_0(X) \approx \beta'X.
\end{equation*}
The best linear predictor (BLP) problem in this case is given by
\begin{equation*}
    \min_{b\in\mathbb{R}^p} \E [(s_o(X) - X'b)^2].
\end{equation*}
We can then make use of the following identification result (which is sometimes referred to as the Horvitz-Thomson transform):
\begin{equation*}
    s_0(Z) = \E[YH|Z]
\end{equation*}
where $H = \frac{D-p(Z)}{p(Z)(1-p(Z))}$.
It is then easy to derive the solution of the BLP problem above as
\begin{equation*}
    \beta = E[XX']^{-1}E[XHY]
\end{equation*}
which instead of the BLP of $Y$ uses the transformed outcome $HY$.
For this we can then make use of standard least squares theory --- given that the nuisance parameter estimation in the first step satisfies certain regularity conditions --- and show
\begin{equation*}
    \hat{\beta} - \beta_0 \sim_a N(0, \Omega/N).
\end{equation*}

@@

@@

% \setstretch{1}
\printbibliography{}
\setstretch{1.5}

>>>>

\title{Topics Causal Inference}

\author{Julian Budde\thanks{Bonn University. Email: s93jbudd[at]uni-bonn.de}}

@@

@@

\paragraph*{Approximations to the CATE}
~\cite{semenova2021debiased} propose to focus on a \textit{low-dimensional} subset $X$ of the covariates $Z$, which the researcher has to decide on ex ante.
Then, the new target for inference becomes the \textit{partial} CATE $s_0(X)$.
Their paper then simultaneously allows for \textit{unconfoundedness} which poses the challenge of estimating $p(Z)$ --- potentially using machine learning methods --- in a first step.
In a second step, they the nfocus on a \textit{linear} approximation to the (partial) CATE\@.
Hence, they essentially use the following approximations to make the problem solvable:
\begin{equation*}
    s_0(Z) \approx s_0(X) \approx \beta'X.
\end{equation*}
The best linear predictor (BLP) problem in this case is given by
\begin{equation*}
    \min_{b\in\mathbb{R}^p} \E [(s_o(X) - X'b)^2].
\end{equation*}
We can then make use of the following identification result (which is sometimes referred to as the Horvitz-Thomson transform):
\begin{equation*}
    s_0(Z) = \E[YH|Z]
\end{equation*}
where $H = \frac{D-p(Z)}{p(Z)(1-p(Z))}$.
It is then easy to derive the solution of the BLP problem above as
\begin{equation*}
    \beta = E[XX']^{-1}E[XHY]
\end{equation*}
which instead of the BLP of $Y$ uses the transformed outcome $HY$.
For this we can then make use of standard least squares theory --- given that the nuisance parameter estimation in the first step satisfies certain regularity conditions --- and show
\begin{equation*}
    \hat{\beta} - \beta_0 \sim_a N(0, \Omega/N).
\end{equation*}

@@

\subsection{Simulation Setup}
\paragraph*{Designing a High-Dimensional DGP}
A key question is how to design a DGP that is both ``realistic'', in the sense that the results are relevant to empirical applications, and still poses a high-dimensional problem.
The typical approach taken in the literature, e.g.\ in~\cite{wager2018estimation} is to postulate a DGP that has convenient analytical features and is easy to simulate.
This simplifies the analysis but begs the question of relevance. Also, it leaves some degree of freedom to the researcher to choose a DGP that favourably displays the advantages of their proposed method.
To address this issue, ``empirical'' methods for simulation design have been proposed, such as the WGAN approach in~\cite{athey2024wgan}.
Without going into detail, the main idea is to train a \textit{generative} machine learning model, that can generate ``new'' observations by estimating the underlying DGP.
The training/optimization procedure can be thought of as a two-player game in which a \textit{generator} and a \textit{discriminator} compete in classifying observations as real or generated.

While this approach can be usefully employed to design specific simulations given a dataset at hand for some questions, it also has some limitations, specifically with respect to the problem studied here.
In particular, we are concerned with a high-dimensional, non-sparse structure of the CATE functional --- this is what the generic ML method is robust against.
However, for a given DGP with high-dimensional structure it cannot be possible train a generator that preserves this structure.
Else, if we had an estimate of the DGP, we could simply construct a plug-in estimator of the CATE, but as argued before such an estimator cannot exist.

For this reason I resort to the analytical DGP presented in~\cite{wager2018estimation} (WA in the following).
To illustrate the difficulty with using a WGAN generated DGP when the problem becomes high-dimensional, I also report results based on a generator that was trained on a draw from the WA dgp.
To the best of my knowledge, such simulation results are not available in the original or any other paper.
\footnote{They do, however, assess robustness of their simulation setup in several dimensions, including sampling variability, in Section 5.}
The following paragraphs explain the simulation setup in more detail.

\paragraph*{WA DGP}
I use the third DGP discussed in~\cite{wager2018estimation} as my baseline DGP\@.
Covariates are generated as $X~U([0, 1]^d)$ and potential outcomes are drawn from $Y^{(0,1)} \sim N(\E[Y^{(0,1)}|X], 1)$.
~\cite{wager2018estimation} refer to $d$ as the ``ambient dimension''. Note that in their simulations this is \textit{not} the dimensionality of the CATE function.
Instead, the CATE function is generated by
\begin{equation*}
    \tau(x) = \xi(X_1)\xi(X_2), \qquad \xi(x) = \frac{2}{1 + \exp^{-12(x-1/2)}}.
\end{equation*}
That is, the CATE function only depends on two covariates and the other $d-2$ covariates are pure noise (in terms of CATE estimation).
The propensity score is kept constant at 0.5.

\paragraph*{WGAN Approximation}
[Describe how WGAN was performed.]

\subsection{Results: WA Replication and WGAN}

\subsection{Results: Generic Machine Learning}

@@

% \setstretch{1}
\printbibliography
\setstretch{1.5}
